---
title: "Data Analysis and Visualization in R - Study Case 1"
output:
  html_document: default
  html_notebook: default
---

### Daniela Simancas, Miguel Sandim, Paula Fortuna
### TUM - Data Analysis and Visualization in R

```{r include=FALSE}
library(data.table)
library(tidyr)
library(dplyr)
library(ggplot2)
library(gridExtra)
```

# 1. Read the data.

```{r}
expression <- fread("expression.txt")
gene <- fread("gene.txt")
genotype <- fread("genotype.txt")
growth <- fread("growth.txt")
marker <- fread("marker.txt")
```

# 2. Tidy the data

```{r}
genotype <- genotype %>% gather("marker", "type", starts_with("mrk_"))
head(genotype)
```

```{r}
expression <- expression %>% mutate(gene = rownames(expression)) %>% gather("stuff", "expression", starts_with("Y")) %>% separate(col = stuff, into = c("environment", "strain"), sep="\\-")
head(expression)
```


```{r}
growth <- growth %>% gather("environment", "growth", starts_with("Y"))
head(growth)
```

# 3. Data pre-processing

```{r}
genotype_type <- genotype %>% group_by(strain) %>% summarise(NLabStrain = table(type)[["Lab strain"]], NWildIsolate = table(type)[["Wild isolate"]])
head(genotype_type)
```

```{r}
growth <- left_join(growth, genotype_type, by = "strain") %>% gather("type", "count", NLabStrain, NWildIsolate)
head(growth)
```

# 4. Does the genotype also show differences in other media?

In order to analyze the growth differences in the several environments, we plotted the growth of the several yeast strains according to number of markers belonging to the Lab Strains and the Wild Isolates.

We concluded that the presence of markers belonging to Lab Strains and Wild Isolates doesn't affect the growth. However, it's ths environment itself that affects this variable (e.g. in YPD_Rapa and YPE the growth is clearly lower than the other environments, independently of the number of markers from each type).

```{r}
ggplot(growth, aes(count, growth)) +
  geom_point() + geom_hex() +
  facet_grid(type~environment)
```

We then plotted the genes with more expression (among all the strains) in the several environments. We concluded that the genes 3080, 3266 1477 and 1511 are the ones that expressed more in all of the environments.

```{r}
mean_expression <- expression %>% group_by(environment, gene) %>% summarise(mean_exp = mean(expression)) %>% arrange(environment, desc(mean_exp)) %>% group_by(environment) %>% top_n(n = 10)

ggplot(mean_expression, aes(x = reorder(factor(gene), mean_exp), y = mean_exp, fill = gene)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  #xlim(5,7) +
  #scale_x_continuous(limits = c(5, 7)) +
  facet_wrap(~environment, scales="free") +
  theme(legend.position="none")
  
```

We match for each marker the respective gene, but then we found out that we only had a very small amount of genes that were marked, and our featuring genes were not contemplated. This way, the chart we ploted was always empty.

In order to solve this, our goal (which we didn't finish) was to try to find the nearest marker for every gene (so we could have information of the markers for the genes we are interested in) This way we could identify the type (lab strain or wild isolate).
